---
title: "covid19 stats"
author: "christophe@pallier.org"
date: "4/20/2020"
output: html_document
---


```{r init}
library(ggplot2)
library(utils)
```


We import "COVID-19 cases worldwide" from https://data.europa.eu/euodp/en/data/dataset/covid-19-coronavirus-data:

```{r import}
covid19ori  <- read.csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv", na.strings = "", fileEncoding = "UTF-8-BOM")
```

We compute the cumulated number of deaths, per countries and per days, and the mortality rates relative to the population (in ppm=parts per million).

```{r newvars}
covid19ori$date <- as.Date(with(covid19ori, sprintf("%4d-%02d-%02d", year, month, day)))
covid19ori <- covid19ori[order(covid19ori$countryterritoryCode, covid19ori$date),]
covid19ori$cumdeaths <- ave(covid19ori$deaths, covid19ori$countryterritoryCode, FUN=cumsum)
covid19ori$cumdeathsnorm <- (covid19ori$cumdeaths / covid19ori$popData2018) * 1e6
covid19ori$deathrateppm <- (covid19ori$deaths / covid19ori$popData2018) * 1e6
```


We select the countries we are interested in.(Ignore the column named 'confined': it is not (yet) used)

```{r selection}
countries <- read.csv(text="code,confined
FRA,1
ITA,1 
ESP,1
BEL,1
GBR,0
DEU,0
NLD,0
SWE,0
USA,NA
DNK,0
NOR,0
FIN,0
", header=TRUE)

covid19 <- subset(covid19ori, countryterritoryCode %in% countries$code & month %in% c(3,4))  # keep only March/April
covid19 <- merge(covid19, countries)

```

And we start to plot the mortality rates:

```{r deathrate, fig.dim=c(10, 10)}

pg <- ggplot(covid19, aes(x=date, y=deathrateppm, group=1)) + geom_line() + geom_point(size=.5) + geom_smooth() + ggtitle("Mortality rates (ppm) by country") + facet_wrap(~countryterritoryCode) + coord_fixed(ratio=1)

print(pg)


```

Same data but overlapping the curves on a single plot. 

```{r deathrate2, fig.dim=c(10, 10)}
pg <- ggplot(covid19, aes(x=date, y=deathrateppm, group=countryterritoryCode, color=countryterritoryCode)) + geom_line(size=1) + geom_point(size=.5) + ggtitle('Mortality rate (ppm) by countries') + coord_fixed(ratio=1) + theme(panel.background = element_rect(fill = "lightgray", size = 2, linetype = "solid"))

print(pg)
```


Next, we plot the cumulative number of the death normalized by country size. 

```{r cumulated_deaths, fig.dim=c(10, 8)}
pg <- ggplot(covid19, aes(x=date, y=cumdeathsnorm, group=countryterritoryCode, color=countryterritoryCode)) + geom_line(size=1) + geom_point(size=.5) + ggtitle('Cumulated deaths (ppm) by countries') + coord_fixed(ratio=0.1) + theme(panel.background = element_rect(fill = "lightgray",size = 2, linetype = "solid"))

print(pg)
```
